### 2021-02-11 ch03_esmvaltool/esmvaltool.tex
Finished chapter 05

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_psi.pdf}
    \caption{}
    \label{fig:03:cox18:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_ec.pdf}
    \caption{}
    \label{fig:03:cox18:b}
  \end{subfigure}
  \caption{Emergent constraint on the \acf{ECS} from global temperature
    variability \autocite{Cox2018}. Magenta colors indicate high sensitivity
    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
    indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
      K^{-1}}$). Both panels have been created with the \acs{ESMValTool} using
    \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
    Temperature variability metric $\Psi$ versus time for historical
    simulations from \acs{CMIP}5 models (colored lines) and observations from
    the HadCRUT4 dataset (black circles) \autocite{Morice2012}. (b) Emergent
    relationship between \acs{ECS} and the temperature variability metric
    $\Psi$. The dot-dashed black line shows the linear regression across the
    climate model ensemble with its associated uncertainties indicated by the
    dashed black lines (standard prediction error; see \cref{eq:02:spe}). Blue
    lines show the observational constraint from HadCRUT4 with its best
    estimate (dot-dashed line) and standard deviation (dashed lines). The
    letters represent individual \acs{CMIP}5 models (see \textcite{Cox2018}
    for details).}
  \label{fig:03:cox18}
\end{figure}

In addition to the ported recipes, several new recipes and diagnostic scripts
have been contributed to the public \ac{ESMValTool} repository (see
\cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
analyses performed in other chapters of this thesis, namely
\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
\emph{recipe\_meehl20sciadv.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
who introduce an emergent constraint in \ac{ECS} based on a global temperature
variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
years for \acs{CMIP}5 models and observations form HadCRUT4
\autocite{Morice2012}. Since high sensitivity climate models exhibit higher
values of $\Psi$ than low sensitivity models, an emergent relationship between
\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.

>>>>

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_psi.pdf}
    \caption{}
    \label{fig:03:cox18:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_ec.pdf}
    \caption{}
    \label{fig:03:cox18:b}
  \end{subfigure}
  \caption{Emergent constraint on the \acf{ECS} from global temperature
    variability \autocite{Cox2018}. Magenta colors indicate high-sensitivity
    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
    indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
      K^{-1}}$). Both panels have been created with the \acs{ESMValTool} using
    \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
    Temperature variability metric $\Psi$ versus time for historical
    simulations from \acs{CMIP}5 models (colored lines) and observations from
    the HadCRUT4 dataset (black circles) \autocite{Morice2012}. (b) Emergent
    relationship between \acs{ECS} and the temperature variability metric
    $\Psi$. The dot-dashed black line shows the linear regression across the
    climate model ensemble with its associated uncertainties indicated by the
    dashed black lines (standard prediction error; see \cref{eq:02:spe}). Blue
    lines show the observational constraint from HadCRUT4 with its best
    estimate (dot-dashed line) and standard deviation (dashed lines). The
    letters represent individual \acs{CMIP}5 models (see \textcite{Cox2018}
    for details).}
  \label{fig:03:cox18}
\end{figure}

In addition to the ported recipes, several new recipes and diagnostic scripts
have been contributed to the public \ac{ESMValTool} repository (see
\cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
analyses performed in other chapters of this thesis, namely
\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
\emph{recipe\_meehl20sciadv.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
who introduce an emergent constraint in \ac{ECS} based on a global temperature
variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
years for \acs{CMIP}5 models and observations form HadCRUT4
\autocite{Morice2012}. Since high-sensitivity climate models exhibit higher
values of $\Psi$ than low sensitivity models, an emergent relationship between
\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.

### 2021-02-11 ch05_paper_ecs/paper_ecs.tex
Finished chapter 05

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
(seasonal) and long (centennial under a forcing) timescales, this quantity is
well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
models. Together with observations of cloud fraction from Cloudsat/CALIPSO
\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011}, and
vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
\ac{MBLC} fraction to changes in \ac{SST} forms an emergent constraint on
\ac{ECS}. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5 ($R^2 = 0.05$) and
$\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 = 0.62$). In their original
publication, \textcite{Zhai2015} find an \ac{ECS} range of
$\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
data (in combination with \acs{CMIP}3 data) has successfully been used in the
original publication, our approach finds that the statistical significance of
the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
shows this extreme behavior (only one other constraint, SHS, shows a slightly
higher significance in \acs{CMIP}6; all other constraints show lower
significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
is the set of climate models used. For our analysis, we use eleven additional
\acs{CMIP}5 models that have not been used in the original publication
(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and MPI-ESM-P). Due to a lack of
publicly available data, the model CESM1-CAM5 that is used in the original
publication is not included in our analysis. The effect of choosing different
subsets of \acs{CMIP}5 models on the emergent relationship is illustrated in
\cref{fig:05:zha}. Using the original \acs{CMIP}5 models from the original
publication gives a considerably higher correlation ($R^2 = 0.38$) than using
all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a strong
dependency of this emergent constraint on the subset of climate models used.
Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
of all the constraints, and much better than on either subset of \acs{CMIP}5
models.

@@

In summary, the $R^2$ of the emergent relationships and the constrained range
of \ac{ECS} each depend strongly on the climate model ensemble used, even
though a physical explanation is given for each emergent constraint that is
thought to be valid for every climate model ensemble. The same behavior is
found for the statistical significance of the emergent relationships using the
null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Except for the ZHA and SHS constraints, every emergent relationship
investigated shows a lower statistical significance (\ie{} a higher $p$- value)
in the \acs{CMIP}6 ensemble than in the \acs{CMIP}5 ensemble. If a conventional
significance test of $p < 0.05$ was applied, eight of the eleven constraints
would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
VOL, and ZHA) would pass this test on \acs{CMIP}6. This is still much better
than would be expected purely by chance. Hence, there is still skill in at
least a few of the constraints, but it is much lower than was suggested in
nearly all of the initial studies.

@@

@@

>>>>

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
(seasonal) and long (centennial under a forcing) time scales, this quantity is
well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
models. Together with observations of cloud fraction from Cloudsat/CALIPSO
\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and
vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
\ac{MBLC} fraction to changes in \ac{SST} forms an emergent constraint on
\ac{ECS}. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5 ($R^2 = 0.05$) and
$\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 = 0.62$). In their original
publication, \textcite{Zhai2015} find an \ac{ECS} range of
$\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
data (in combination with \acs{CMIP}3 data) has successfully been used in the
original publication, our approach finds that the statistical significance of
the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
shows this extreme behavior (only one other constraint, SHS, shows a slightly
higher significance in \acs{CMIP}6; all other constraints show lower
significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
is the set of climate models used. For our analysis, we use eleven additional
\acs{CMIP}5 models that have not been used in the original publication
(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and MPI-ESM-P). Due to a lack of
publicly available data, the model CESM1-CAM5 that is used in the original
publication is not included in our analysis. The effect of choosing different
subsets of \acs{CMIP}5 models on the emergent relationship is illustrated in
\cref{fig:05:zha}. Using the original \acs{CMIP}5 models from the original
publication gives a considerably higher correlation ($R^2 = 0.38$) than using
all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a strong
dependency of this emergent constraint on the subset of climate models used.
Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
of all the constraints and much better than on either subset of \acs{CMIP}5
models.

@@

In summary, the $R^2$ of the emergent relationships and the constrained range
of \ac{ECS} each depend strongly on the climate model ensemble used, even
though a physical explanation is given for each emergent constraint that is
thought to be valid for every climate model ensemble. The same behavior is
found for the statistical significance of the emergent relationships using the
null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Except for the ZHA and SHS constraints, every emergent relationship
investigated shows a lower statistical significance (\ie{} a higher $p$- value)
in the \acs{CMIP}6 ensemble than in the \acs{CMIP}5 ensemble. If a conventional
significance test of $p < 0.05$ was applied, eight of the eleven constraints
would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
VOL and ZHA) would pass this test on \acs{CMIP}6. This is still much better
than would be expected purely by chance. Hence, there is still skill in at
least a few of the constraints, but it is much lower than suggested in nearly
all of the initial studies.

@@

As shown in the previous sections, most emergent relationships show smaller
coefficients of determination when evaluated on the new \acs{CMIP}6 ensemble
compared to the \acs{CMIP}5 ensemble. In this section, we discuss possible
reasons for, and implications of, these differences. As reported by
\textcite{Caldwell2014}, the large amount of data provided by modern \acp{ESM}
can generate spurious correlations of variables between past climate and
\ac{ECS} just by chance, especially when only a small number of climate models
is considered. This would cause the performance of the emergent constraint to
be reduced on out-of-sample data (like the new \acs{CMIP}6 ensemble), since the
emergent relationship appeared just by chance and not because of a physically
based mechanism.

A further reason for the weaker emergent relationships in \acs{CMIP}6 may be
the increased complexity of the participating \acp{ESM}. Each emergent
constraint approach is based on the assumption that a single observable process
or physical aspect in the current climate dominates the uncertainty in
\ac{ECS}. Some emergent constraints such as ZHA and BRI relate changes in cloud
properties (ZHA: low-level cloud fraction; BRI: cloud reflectivity) on seasonal
or interannual time scales driven by changes in \ac{SST} to \ac{ECS}. This
means that it has to be implicitly assumed that the observable changes in these
properties on seasonal or interannual time scales are related to those
occurring as a result of climate forcing in a way successfully captured by the
\acp{ESM}. While this assumption seems to make sense, we do not know whether
the \acp{ESM} cover all relevant processes of the real Earth system. For
example, it may be possible that there exist processes that are unimportant in
the \acp{ESM} (and hence are not captured by the emergent constraints) but are
actually important in reality. This lack of relevant processes may lead to an
overconfident constraint. Thus, the more complex \acp{ESM} of the \acs{CMIP}6
ensemble are more likely to capture relevant processes of the true climate,
which leads to weaker emergent relationships. On the other hand, emergent
constraints on the less complex \acs{CMIP}3 and \acs{CMIP}5 ensemble may be
overconfident.

For \acs{CMIP}6 models, \textcite{Zelinka2020} show that cloud feedbacks and
thus \ac{ECS} in high-sensitivity models are to some extent associated with
changes in clouds over the Southern Ocean, while in \acs{CMIP}3 and \acs{CMIP}5
the uncertainty in cloud feedbacks is dominated by clouds in the subtropical
subsidence regions. One might speculate that a possible reason for this might
be an improved simulation of clouds over the Southern Ocean in some models
\autocite{BodasSalcedo2019, Gettelman2019}, as shown for some pre-\acs{CMIP}6
model versions evaluated by \textcite{Lauer2018}. The findings of
\textcite{Zelinka2020} can also at least partly explain the larger inter-model
spread in climate sensitivity due to a greater diversity of cloud feedbacks,
which also results in a weaker emergent constraint compared with \acs{CMIP}5
models, as most of them constrained low-cloud feedbacks. They find that on
average, the shortwave low-cloud feedback is larger in \acs{CMIP}6 than in
\acs{CMIP}5, which they primarily relate to changes in the representation of
clouds. As a possible explanation, \textcite{Zelinka2020} give an increase in
mean-state supercooled liquid water (\ie{} increase in the cloud water liquid
fraction) in mixed-phase clouds resulting in less pronounced increases in
low-level cloud cover and water content with warmer \acp{SST} particularly in
midlatitudes.

Our findings suggest that the process-oriented emergent constraints (\ie{} all
of the emergent constraints investigated here except COX) are only successful
in constraining \ac{ECS} as long as the uncertainty in \ac{ECS} is dominated by
the same process or feedback. In the \acs{CMIP}5 ensemble, cloud feedback is
the main contributor to the spread in \ac{ECS} with low-level clouds in
tropical subsidence regions dominating the spread in cloud feedback
\autocite{Ceppi2017}. If any other process or feedback is biased (or missing)
in the ensemble as a whole, then these process-oriented emergent constraints
will be biased in their estimates of \ac{ECS}. The appearance of diverse new
feedback processes in \acs{CMIP}6 could explain the reduced skill when applied
to \acs{CMIP}6 data, and a tendency for these to be positive would explain the
upward shift in the model \ac{ECS} distribution that is not captured by the
\acs{CMIP}5-trained constraints. Process-oriented emergent constraints are
therefore perhaps best thought of as constraints on the processes that they
target, rather than constraints on \ac{ECS}.

Emergent constraints that use global temperature change as a way of
constraining \ac{ECS} could in principle overcome this problem. If one feedback
is biased in an ensemble the constraint might still work as both global
temperature change and \ac{ECS} might similarly reflect the sum of all
feedbacks. Emergent constraints of this kind include the tropical temperature
during the Last Glacial Maximum \autocite{Hargreaves2012}, tropical temperature
anomalies during the mid-Pliocene Warm Period \autocite{Hargreaves2016} and
post-1970s warming \autocite{JimenezdelaCuesta2019}. This seems to be supported
by the findings of \textcite{Tokarska2020}, who test an emergent constraint for
the \ac{TCR} based on recent global warming trends on the \acs{CMIP}5 and
\acs{CMIP}6 ensembles with similar results for both model ensembles.
Nevertheless, these temperature-based estimates are sensitive to assumptions
about forcings and unforced decadal temperature variations, which could also be
incorrect, as could model-predicted relationships between feedbacks on short
and long time scales that are implicit in most such measures.

@@

This paper assesses eleven different emergent constraints on \ac{ECS}, of which
most are directly or indirectly related to cloud feedbacks, by applying them to
results from \acp{ESM} contributing to \acs{CMIP}5 and \acs{CMIP}6. Of
particular interest are the results from \acs{CMIP}6, since all analyzed
emergent constraints have been published prior to the availability of
\acs{CMIP}6 data. In summary, the best estimate of \ac{ECS} averaged over all
emergent constraints increases by $12 \unit{\%}$ when moving from relationships
trained on \acs{CMIP}5 to those trained on \acs{CMIP}6. Some increase is
predicted by almost every constraint we analyze (with TII as the only
exception) and can be at least partly explained by the increased multi-model
mean \ac{ECS} of \acs{CMIP}6, which has not been accompanied by systematic
changes in the constraint variables that could explain this increase, leading
to regression fits with higher intercept values at observed constraint values.
This is also illustrated by the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM} in the
left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

Our results also show that, except for ZHA, all emergent relationships are
weaker (in terms of the coefficient of determination $R^2$) in \acs{CMIP}6
compared to \acs{CMIP}5, which means that the corresponding emergent
relationships are able to explain less of the \ac{ECS} variation simulated by
the newer \acs{CMIP}6 models than by those of \acs{CMIP}5. This is also
demonstrated by the statistical significance, which is lower in \acs{CMIP}6
than in \acs{CMIP}5 for all emergent constraint except for ZHA and SHS. As
described in \cref{subsec:02:emergent_constraints}, our test for statistical
significance uses the null hypothesis that there is no correlation between the
predictor variable of the emergent constraint and \ac{ECS}. Further evidence of
the decreased performance of the emergent constraints in \acs{CMIP}6 is given
by the size of the constrained \acs{ECS} ranges, which widens by $51 \unit{\%}$
in \acs{CMIP}6 compared to \acs{CMIP}5 on average. Moreover, our more detailed
analysis of the ZHA constraint (see \cref{fig:05:zha}) shows that this emergent
constraint is very sensitive to outliers and the subset of the climate model
ensemble used to fit the emergent relationship. Such a behavior might not be
unique to the ZHA constraint but could apply to other emergent constraints as
well. This in turn suggests that the number of climate models commonly used for
emergent constraints might be too low, leading to non-robust relationships.

Our analysis makes a number of simplifying assumptions common to other studies,
such as model independence, discussed in detail in
\cref{subsec:02:ecs,subsec:02:emergent_constraints}. These assumptions affect
the significance of emergent relationships and the \acp{PDF} of \ac{ECS} based
on a constraint. However, they do not affect our main conclusions here, which
concern the change in performance on \acs{CMIP}6 relative to \acs{CMIP}5 and
the implications for robustness and future use of emergent constraints.

\ac{ECS} is the product of the complex interactions of the many components and
feedbacks. Thus, constraining \ac{ECS} with a single physical process might
overly simplify this problem. Such single process-oriented emergent constraints
therefore do not seem to be helpful in constraining \ac{ECS} but should
probably rather be thought of as constraints for the process or feedback they
are actually targeting (if that can be clearly identified). With increasing
computational resources available to climate science, more and more detailed
process interactions can be taken into account in a modern \ac{ESM}. In
contrast, the predecessor versions \acs{CMIP}3 and \acs{CMIP}5 are less complex
with simpler atmospheric process representations, so constraining uncertainties
of a single dominant process may allow for an apparently more successful
constraint of \ac{ECS} than would be achieved in more complex models. As a
conclusion, we argue that to constrain \ac{ECS} in a more robust way, it might
be beneficial to apply multivariate approaches that are able to consider
multiple (different) relevant physical processes and feedbacks at once and thus
are able to get a broader picture of the complex reality. A possible approach
for this is given by \textcite{Bretherton2020}, who combine the information
from multiple emergent constraints on \ac{ECS} using a multivariate Gaussian
and multiple linear regression. For the \acs{CMIP}3 and \acs{CMIP}5 ensembles,
they find an increased best estimate relative to the unweighted ensemble mean
similar to the participating individual emergent constraints, but with lower
uncertainty range. Moreover, new machine learning techniques are a promising
avenue forward for such multivariate approaches and for constraining
uncertainties in multi-model projections \commentcite{Schlund2020}{see
  \cref{ch:06:paper_gpp}} with the aim of further improving climate modeling
and analysis \autocite{Reichstein2019}.
